Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-09-03 draft.tex
Revised Abstract with Sylvia

% \abstract{
% The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words. The text must be set to 9-point font size. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.}
\abstract{ 
% Motivation
Epidemiology is the study of causation of diseases.
%%
Subject groups (a cohort) is analyzed to correlate lifestyle and medical variables with diseases.
%%
Most recently, large cohort studies also acquire medical image data.
%% Problem statement
We correlate image-derived variables from the lumbar spine with lower back pain.
%% Approach
Our analysis is based on detection masks from a semi-automatic model based detection of 2,540 lumbar spines.
%%
From the detection masks, we derive shape-describing variables, such as course, curvature and torsion of the spine canal.
%%
We classify non-image parameter using decision trees created using the C4.5/C5.0 algorithm.
%%
The analysis yields a decision tree per non-image parameter.
%%
We make these trees comparable by introducing tree size and a classification error rate metrics.
%%
A interactive scatterplot visualization allows to compare the decision trees.
%%
%\\\\
% Results
% Conclusions
We conclude that the metrics extracted from the detection masks are only sufficient for classification of lifestyle factors, such as gender or nutrition.
%%
The presented approach is a valuable extension to the interactive visual analysis of epidemiological data and not limited to image-parameter analysis.
}

>>>>

% \abstract{
% The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words. The text must be set to 9-point font size. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.}
\abstract{ 
The goal of epidemiology is to provide insight into disease causations.
%%
Hence, subject groups (cohorts) are analyzed to correlate the subjects' varying lifestyles and medical properties with diseases.
%%
Recently, these cohort studies comprise medical image data as well.
%%
We assess the predictive quality of image-derived variables for lower back pain.
%%
The evaluation is embedded in an interactive visual analysis (IVA) framework that is based on 2,540 segmented lumbar spine data sets.
%%
The segmentation results are evaluated and quantified by employing shape-describing variables, such as the spine canal course, its curvature and torsion.
%%
We analyze the predictive power of shape-describing variables for non-image variables, e.g., pain indicators.
%%
Therefore, we automatically extract a decision tree classifier for each non-image variable.
%%
Furthermore, we provide an IVA technique to compare the classifiers with interactive scatterplot visualizations.
%%
As a first result, we conclude that image-based variables are only sufficient for classification of lifestyle factors.
%%
A correlation between lumbar spine shape and lower back pain disease was not existent in the automatically trained classifiers.
%%
However, the presented approach is a valuable extension for the IVA of epidemiological data.
%%
Hence, correlations between non-image parameters were successfully detected and characterized.
}
